### 2024-04-09 latex/draft.tex
Pushing initial latex creation.

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}

\title{Causal Inference with Graphical Neural Networks}
\author{Your Name}
\date{\today}
\maketitle

\section{Introduction}

% Introduce the topic of causal inference with Graphical Neural Networks (GNNs).

\section{Background}

% Provide background information on causal inference, neural networks, and Graphical Neural Networks (GNNs).

\section{Methodology}

% Describe the methodology of using GNNs for causal inference. Include information on how GNNs are trained, how causal inference is performed, etc.

\section{Experiments}

% Present the experiments conducted to evaluate the performance of GNNs for causal inference. Include dataset descriptions, experimental setup, and results.

\subsection{Dataset Description}

% Describe the datasets used in the experiments.

\subsection{Experimental Setup}

% Explain the experimental setup, including model architecture, training procedure, hyperparameters, etc.

\subsection{Results}

% Present and analyze the results of the experiments. Include metrics, comparisons with baseline methods, and any insights gained.

\section{Discussion}

% Discuss the implications of the results, limitations of the approach, potential future research directions, etc.

\section{Conclusion}

% Summarize the key findings of the study and conclude the paper.

\section*{Acknowledgments}

% Acknowledge any individuals or organizations that contributed to the project.

% Bibliography
\bibliographystyle{plain}
\bibliography{references} % Replace 'references' with the name of your .bib file

\end{document}

### 2024-04-09 latex/draft.tex
Pushing some latex changes.

@@
 \usepackage{graphicx}
+\usepackage{ntheorem}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
 
@@
 % Introduce the topic of causal inference with Graphical Neural Networks (GNNs).
+
+\section{$\textbf{do}$ Operator}
+
+The \textbf{do} operator is a way to represent interventions in a causal model. 
+It is a way to represent the effect of an intervention on a variable. As an example, 
+consider the following model involving smoking.
+
+If a person's fingernails $(N)$ have turned yellow, this implies a higher probability 
+that they are a heavy smoker $(S)$ and hence have a higher probability of developing 
+lung cancer $(C)$. But, simply dyeing a persons fingernails yellow does not impact 
+their probability of developing lung cancer.
+
+So, in terms of $\textbf{do}$ calculus, we can denote the process of setting a 
+variable $N$ to have a value $\textit{yellow}$ by $\textbf{do}(N = \textit{yellow})$. 
+We note that
+
+\begin{equation*}
+P(C \;|\;N = \textit{yellow}) \neq P(C\;|\; \textbf{do}(N=\textit{yellow})).
+\end{equation*}
+
+With this in mind, we now define the $\textbf{do}$ operator.
+
+\begin{theorem}[{{\cite{pearl2009causal}}}]
+    In a causal diagram $\Gamma$ with nodes $X_1,\dots, X_n$ and joint distribution 
+    $P(X_1, \dots, X_n)$, the result of doing $X_i = x_i$ on the joint distribution is
+
+    \[
+        P(X_1, \dots, X_n \;|\; \textbf{do}(X_i = x_i)) = \frac{P(x_1,\dots,x_n)}{P(x_i\;|\; \textup{par}(x_i))} = \prod_{j\neq i}P(x_j\;|\; \textup{par}(x_j)).
+    \]
+\end{theorem}
+
+In this, we have $\text{par}(x_i)$ represent values of the parent nodes of $\text{PAR}(X_i)$ of $X_i$ in $\Gamma$.
+The probabilities on the right hand side of the above equation are what we call \textit{preintervention}. 
+This means they use the original probabilities from the original model before doing $X_i = x_i$.
+
+It is important to note that the above equation is how we calculate the probability of several events 
+happening given one event has happened. What if we want to get the probability of a single event happening, 
+given we do a single event? That leads to the following corollary.
+
+\begin{corollary}
+    If $X$ and $Y$ are random variables in a causal diagram $\Gamma$ and $\textup{PAR}(X)$ are the parents of $X$, then
+
+    \[P(y\;|\;  \textbf{do}(x)) = \sum_{\textup{par}}\frac{P(x,\,y,\,\textup{par})}{P(x\;|\; \textup{par})},\]
+
+    where the sum runs over all values $\textup{par}$ that the variables $\textup{PAR}(X)$ can take. If $X$ has no parents, then
+
+    \[P(y\;|\; \textbf{do}(x)) = \frac{P(x,\,y)}{P(x)} = P(y\;|\; x).\]
+\end{corollary}
 

### 2024-04-09 latex/draft.tex
Pushing latex. Still working on the DAG in figure 2.

@@
 \newtheorem{lemma}[theorem]{Lemma}
+\usepackage{tkz-graph}
+\usetikzlibrary{arrows}
+\usetikzlibrary{shapes.geometric}
+
+\SetVertexNormal[Shape      = circle,
+                 FillColor  = cyan,
+                 LineWidth  = 2pt]
+\SetUpEdge[lw         = 1pt,
+           color      = black,
+           labelcolor = white,
+           labeltext  = red,
+           labelstyle = {sloped,draw,text=blue}]
+\tikzset{Bullet/.style={fill=black,draw,color=#1,circle,minimum size=3pt,scale=0.75}}
 
@@
     \]
+    \label{theorem:do}
 \end{theorem}
@@
     \[P(y\;|\; \textbf{do}(x)) = \frac{P(x,\,y)}{P(x)} = P(y\;|\; x).\]
+    \label{corollary:do}
 \end{corollary}
+
+Let us now consider a basic example to see how this works. Consider the following causal diagram in Figure~\ref{fig:causal_dia}.
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[scale=1.5, Bullet/.style={circle, draw=black, fill=#1, inner sep=0pt, minimum size=3mm, line width=1.pt}]
+        \node[Bullet=cyan,label=above :{$A$}] (A) at (0,2){};
+        \node[Bullet=cyan,label=above :{$C$}] (C) at (4,2){};
+        \node[Bullet=cyan,label=below :{$B$}] (B) at (2,1){};
+        \node[Bullet=cyan,label=below :{$Y$}] (Y) at (4,0){};
+        \node[Bullet=cyan,label=below :{$X$}] (X) at (0,0){};
+
+        \draw[->, line width=1] (A) -- (B);
+        \draw[->, line width=1] (C) -- (B);
+        \draw[->, line width=1] (B) -- (X);
+        \draw[->, line width=1] (C) -- (Y);
+        \draw[->, line width=1] (A) -- (X);
+    \end{tikzpicture}
+    \caption{Basic causal diagram. Note it is in the form of a directed acyclic graph (DAG).}
+    \label{fig:causal_dia}
+\end{figure}
+
+In this diagram, we can see that $A$ and $C$ are both parents of $B$. So, for any values of $x$ and $b$, 
+Corollary~\ref{corollary:do} tells us that
+
+\begin{align*}
+    P(X = x \,|\, \textbf{do}(B=b)) &= \sum_{\text{par}(b)}\frac{P(x, \,b, \,\text{par}(b))}{P(b\,|\, \text{par}(b))} \\
+    %  &= \sum_{a}\sum_{c}\frac{P(X=x\,, \, A=a\,,\, B=b \, ,\, C=c)}{P(B=b \;|\; A=a,\, C=c)} \\
+    %  &= \sum_{a}\sum_{c}\frac{P(X=x\;|\; A=a\,,\, B=b)P(B=b\;|\; A=a,\, C=c)P(A=a)P(C=c)}{P(B=b \;|\; A=a,\, C=c)} \\
+    %  &= \sum_{a}\sum_{c}P(X=x\;|\; A=a\,,\, B=b)P(A=a)P(C=c) \\
+\end{align*}
+
+which, written out, is
+
+\begin{equation*}
+    \sum_{\text{par}(b)}\frac{P(x, \,b, \,\text{par}(b))}{P(b\,|\, \text{par}(b))} = \sum_{a}\sum_{c}\frac{P(X=x\,, \, A=a\,,\, B=b \, ,\, C=c)}{P(B=b \;|\; A=a,\, C=c)}.
+\end{equation*}
+
+By dependence of nodes only on their parents and the rules of probability, this turns into
+
+\begin{equation*}
+    \sum_{a}\sum_{c}\frac{P(X=x\;|\; A=a\,,\, B=b)P(B=b\;|\; A=a,\, C=c)P(A=a)P(C=c)}{P(B=b \;|\; A=a,\, C=c)},
+\end{equation*}
+
+which simplifies to
+
+\begin{equation*}
+    \sum_{a}\sum_{c}P(X=x\;|\; A=a\,,\, B=b)P(A=a)P(C=c).
+\end{equation*}
+
+Since there is only one instance where we are considering the probability with respect to $c$, 
+we can simplify this to
+
+\begin{equation*}
+    \sum_{a}P(X=x\;|\; A=a\,,\, B=b)P(A=a),
+\end{equation*}
+
+which is our final answer.
+
+While this introduction to the \textbf{do} operator might feel a bit abstract, it is the foundation
+of all current research in causal inference.
+
+\section{Data}
+
+For our project, we used the LUCAS0 dataset~\cite{lucas_dataset}, which is a toy data set 
+generated artificially by causal Bayesian networks with binary variables. The LUCAS0 dataset
+is a DAG with 11 nodes and 2000 training samples, where the DAG is represented as
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[scale=1.5, Bullet/.style={circle, draw=black, fill=#1, inner sep=0pt, minimum size=3mm, line width=1.pt}]
+        \draw[help lines, gray] (0,0) grid (6,6);
+        \node[ellipse, draw=black, fill=lime] (A) at (2,2){Anxiety};
+        \node[ellipse, draw=black, fill=lime] (P) at (5,2){Peer Pressure};
+        \node[ellipse, draw=black, fill=teal] (S) at (3,1){Smoking};
+        \node[ellipse, draw=black, fill=lime] (Y) at (1,1){Yellow Fingers};
+        \node[Bullet=cyan,label=below :{$Y$}] (Y) at (4,0){};
+        \node[Bullet=cyan,label=below :{$X$}] (X) at (0,0){};
+
+        \draw[->, line width=1] (A) -- (S);
+        \draw[->, line width=1] (P) -- (S);
+        \draw[->, line width=1] (B) -- (X);
+        \draw[->, line width=1] (C) -- (Y);
+        \draw[->, line width=1] (A) -- (X);
+    \end{tikzpicture}
+    \caption{Basic causal diagram. Note it is in the form of a directed acyclic graph (DAG).}
+    \label{fig:lucas_dag}
+\end{figure}
 

### 2024-04-10 latex/draft.tex
Pushing latex changes.

@@
         \node[ellipse, draw=black, fill=lime] (A) at (2,2){Anxiety};
-        \node[ellipse, draw=black, fill=lime] (P) at (5,2){Peer Pressure};
+        \node[ellipse, draw=black, fill=lime, align=center] (P) at (4,2){Peer\\Pressure};
         \node[ellipse, draw=black, fill=teal] (S) at (3,1){Smoking};
-        \node[ellipse, draw=black, fill=lime] (Y) at (1,1){Yellow Fingers};
-        \node[Bullet=cyan,label=below :{$Y$}] (Y) at (4,0){};
+        \node[ellipse, draw=black, fill=lime, align=center] (Y) at (1,1){Yellow\\Fingers};
+        \node[ellipse, draw=black, fill=lightgray, align=center] (B) at (6, 2){Born on\\Even Day};
+        \node[ellipse, draw=black, fill=teal, align=center] (G) at (5, 1){Genetics};
+        \node[ellipse, draw=black, fill=teal, align=center] (C) at (2, 0){Allergy};
         \node[Bullet=cyan,label=below :{$X$}] (X) at (0,0){};
@@
         \draw[->, line width=1] (P) -- (S);
-        \draw[->, line width=1] (B) -- (X);
-        \draw[->, line width=1] (C) -- (Y);
-        \draw[->, line width=1] (A) -- (X);
+        \draw[->, line width=1] (S) -- (Y);
+
     \end{tikzpicture}

### 2024-04-10 latex/draft.tex
Pushing some fun latex changes.

@@
 \title{Causal Inference with Graphical Neural Networks}
-\author{Your Name}
+\author{Gwen Johnson\and Dylan Skinner\and Dallin Stewart\and Jason Vasquez}
 \date{\today}
@@
 \section{Introduction}
+
+\section{Graph Neural Networks}
+
+Originally proposed in 2005 by Mori et al.~\cite{gori2005new}, Graph Neural Networks 
+(GNNs) are a class of neural networks that operate on graph-structured data. 
+They have gained significant attention in recent years due to their versatility 
+in handling various types of graph data, including social networks, citation networks, 
+biological networks, and more.
+
+Unlike traditional neural networks, which operate on 
+grid-like structures such as images or sequences, GNNs are specifically designed to 
+capture and leverage the structural information present in graphs. One of their 
+key strengths lies in their ability to learn meaningful representations of nodes in a 
+graph, which can then be used for various downstream tasks such as node classification, 
+link prediction, and graph classification. Their ability to capture and model complex 
+relationships in graph data makes them invaluable tools for exploring and understanding 
+real-world phenomena represented in graph form.
+
+In our project, we are utilizing the power of GNNs to perform a supervised classification
+task. This means we are training a GNN on a labeled, tabular dataset, where each graph instance
+(or node) is associated with a target label.
 
@@
 generated artificially by causal Bayesian networks with binary variables. The LUCAS0 dataset
-is a DAG with 11 nodes and 2000 training samples, where the DAG is represented as
+is a DAG with 11 nodes and 2000 training samples, where the DAG is represented as in Figure~\ref{fig:lucas_dag}.
 
@@
     \begin{tikzpicture}[scale=1.5, Bullet/.style={circle, draw=black, fill=#1, inner sep=0pt, minimum size=3mm, line width=1.pt}]
-        \draw[help lines, gray] (0,0) grid (6,6);
+        % \draw[help lines, gray] (0,0) grid (6,6);
         \node[ellipse, draw=black, fill=lime] (A) at (2,2){Anxiety};
@@
         \node[ellipse, draw=black, fill=teal, align=center] (G) at (5, 1){Genetics};
-        \node[ellipse, draw=black, fill=teal, align=center] (C) at (2, 0){Allergy};
-        \node[Bullet=cyan,label=below :{$X$}] (X) at (0,0){};
+        \node[ellipse, draw=black, fill=teal, align=center] (Al) at (2, 0){Allergy};
+        \node[ellipse, draw=black, fill=pink, align=center] (L) at (4, 0){Lung\\Cancer};
+        \node[ellipse, draw=black, fill=lime, align=center] (AD) at (6, 0){Attention\\Disorder};
+        \node[ellipse, draw=black, fill=teal, align=center] (C) at (3, -1){Coughing};
+        \node[ellipse, draw=black, fill=teal, align=center] (F) at (5, -1){Fatigue};
+        \node[ellipse, draw=black, fill=lime, align=center] (CA) at (6, -2){Car\\Accident};
 
@@
         \draw[->, line width=1] (S) -- (Y);
+        \draw[->, line width=1] (S) -- (L);
+        \draw[->, line width=1] (G) -- (L);
+        \draw[->, line width=1] (G) -- (AD);
+        \draw[->, line width=1] (Al) -- (C);
+        \draw[->, line width=1] (L) -- (C);
+        \draw[->, line width=1] (L) -- (F);
+        \draw[->, line width=1] (C) -- (F);
+        \draw[->, line width=1] (F) -- (CA);
+        \draw[->, line width=1] (AD) -- (CA);
 
     \end{tikzpicture}
-    \caption{Basic causal diagram. Note it is in the form of a directed acyclic graph (DAG).}
+    \caption{Basic causal diagram. Note it is in the form of a directed acyclic graph (DAG). Our target variable
+    is shaded in ${\color{pink} \textbf{pink}}$, and the nodes in {\color{teal} \textbf{teal}} constitude the Markov
+    blanket of the target variable.}
     \label{fig:lucas_dag}
@@
 
+\newpage
 % Bibliography

### 2024-04-11 latex/draft.tex
Pushing changes to latex file.

@@
 \usetikzlibrary{shapes.geometric}
+\usepackage{hyperref}
 
@@
 \section{Introduction}
+\label{intro}
 
 \section{Graph Neural Networks}
+\label{gnn}
 
@@
 \section{$\textbf{do}$ Operator}
+\label{do}
 
@@
 \section{Data}
+\label{data}
 
@@
 
+Each node of the graph was is associated with a specific conditional probability that the
+creators of the dataset used to generate the data. These probabilities can be found in Table~\ref{tab:lucas_probs}
+in Appendix~\ref{create}.
+
 \section{Background}
@@
 
+\newpage
+
+\section*{Appendix}
+
+\appendix
+\section{Creating the LUCAS0 Dataset}
+\label{create}
+
+As described in the \nameref{data} section, the LUCAS0 dataset is a toy dataset generated 
+artificially by causal Bayesian networks with binary variables. The dataset consists of 
+11 nodes and 2000 training samples. The causal diagram for the LUCAS0 dataset is shown in Figure~\ref{fig:lucas_dag}.
+
+On the website, the authors mention that each node of the graph is associated with conditional probabilities
+which were used to generate the data. These probabilities are found in Table~\ref{tab:lucas_probs}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{|c|c|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{Conditional Probabilities}} \\
+        \hline
+        \textit{Conditional} & \textit{Probability} \\
+        \hline
+        P(Anxiety=T) & 0.64277 \\ \hline
+        P(Peer Pressure=T) & 0.32997 \\ \hline
+        P(Smoking=T|Peer Pressure=F, Anxiety=F) & 0.43118 \\ \hline
+        P(Smoking=T|Peer Pressure=T, Anxiety=F) & 0.74591 \\ \hline
+        P(Smoking=T|Peer Pressure=F, Anxiety=T) & 0.8686 \\ \hline
+        P(Smoking=T|Peer Pressure=T, Anxiety=T) & 0.91576 \\ \hline
+        P(Yellow Fingers=T|Smoking=F) & 0.23119 \\ \hline
+        P(Yellow Fingers=T|Smoking=T) & 0.95372 \\ \hline
+        P(Genetics=T) & 0.15953 \\ \hline
+        P(Lung cancer=T|Genetics=F, Smoking=F) & 0.23146 \\ \hline
+        P(Lung cancer=T|Genetics=T, Smoking=F) & 0.86996 \\ \hline
+        P(Lung cancer=T|Genetics=F, Smoking=T) & 0.83934 \\ \hline
+        P(Lung cancer=T|Genetics=T, Smoking=T) & 0.99351 \\ \hline
+        P(Attention Disorder=T|Genetics=F) & 0.28956 \\ \hline
+        P(Attention Disorder=T|Genetics=T) & 0.68706 \\ \hline
+        P(Born an Even Day=T) & 0.5 \\ \hline
+        P(Allergy=T) & 0.32841 \\ \hline
+        P(Coughing=T|Allergy=F, Lung cancer=F) & 0.1347 \\ \hline
+        P(Coughing=T|Allergy=T, Lung cancer=F) & 0.64592 \\ \hline
+        P(Coughing=T|Allergy=F, Lung cancer=T) & 0.7664 \\ \hline
+        P(Coughing=T|Allergy=T, Lung cancer=T) & 0.99947 \\ \hline
+        P(Fatigue=T|Lung cancer=F, Coughing=F) & 0.35212 \\ \hline
+        P(Fatigue=T|Lung cancer=T, Coughing=F) & 0.56514 \\ \hline
+        P(Fatigue=T|Lung cancer=F, Coughing=T) & 0.80016 \\ \hline
+        P(Fatigue=T|Lung cancer=T, Coughing=T) & 0.89589 \\ \hline
+        P(Car Accident=T|Attention Disorder=F, Fatigue=F) & 0.2274 \\ \hline
+        P(Car Accident=T|Attention Disorder=T, Fatigue=F) & 0.779 \\ \hline
+        P(Car Accident=T|Attention Disorder=F, Fatigue=T) & 0.78861 \\ \hline
+        P(Car Accident=T|Attention Disorder=T, Fatigue=T) & 0.97169 \\ \hline
+        \hline
+    \end{tabular}
+    \caption{Conditional probabilities for the nodes in the LUCAS0 dataset.}
+    \label{tab:lucas_probs}
+\end{table}
+
 \end{document}

### 2024-04-12 latex/draft.tex
Pushing latex changes.

--- 2024-04-10 latex/draft.tex
@@
 
-\section{Background}
+Each node of the graph was is associated with a specific conditional probability that the
+creators of the dataset used to generate the data. These probabilities can be found in Table~\ref{tab:lucas_probs}
+in Appendix~\ref{create}.
+
+\section{do Calculus Result stuff (idk where this goes)}
+
+Let's consider the following conditional probabilities with \textbf{do} operators applied:
+
+\begin{align*}
+    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{YF} = \text{T})), \\
+    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{PP} = \text{T})), \\
+    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{A} = \text{T})), \\
+    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{AD} = \text{T})), \\
+    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{CA} = \text{T})), \\
+\end{align*}
+
+where LC is lung cancer, YF is yellow fingers, PP is peer pressure, A is anxiety, AD is attention disorder, and CA 
+is car accident. We note that we chose this probability because it does not involve any
+variables that are in the Markov blanket of our target variable. Thus, we should expect that there is very little
+predictive power in this probability.
+
+Plugging forward with \textbf{do} calculus, we can write $P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T}))$ as
+(using corollary~\ref{corollary:do})
+
+\begin{equation*}
+    P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T})) = 
+    \sum_{s\in\{T, F\}} \frac{P(\text{YF} = \text{T},\, \text{LC} = \text{T},\, \text{S} = s)}{P(\text{YF} = \text{T}\,|\, \text{S} = s)},
+\end{equation*}
+
+where $\text{S}$ is smoking. Using the property that we can split the numerator into probabilities only involving parents ({\color{red}rewrite})
+
+\begin{equation*}
+    \sum_{s\in\{T, F\}}\frac{P(\text{YF} =\text{T}\;|\; \text{S} = s) P(\text{LC} 
+    = \text{T}\;|\; \text{S}=s,\, \text{G}) P(\text{S}=s\;|\;\text{A},\, \text{PP})}
+    {P(\text{YF} = \text{T}\;|\; \text{S} = s)}.
+\end{equation*}
+
+All of these fun conditionals lead us to
+
+\begin{equation*}
+    \sum_{s,a,g,p}\frac{P(\text{YF} =\text{T}\;|\; \text{S} = s) P(\text{LC} 
+    = \text{T}\;|\; \text{S}=s,\, \text{G}=g) P(\text{S}=s\;|\;\text{A}=a,\, \text{PP}=p)}
+    {P(\text{YF} = \text{T}\;|\; \text{S} = s)},
+\end{equation*}
+
+where each $s,a,g,p$ is summed over the set $\{T, F\}$.
 
@@
 
+\newpage
+
+\section*{Appendix}
+
+\appendix
+\section{Creating the LUCAS0 Dataset}
+\label{create}
+
+As described in the \nameref{data} section, the LUCAS0 dataset is a toy dataset generated 
+artificially by causal Bayesian networks with binary variables. The dataset consists of 
+11 nodes and 2000 training samples. The causal diagram for the LUCAS0 dataset is shown in Figure~\ref{fig:lucas_dag}.
+
+On the website, the authors mention that each node of the graph is associated with conditional probabilities
+which were used to generate the data. These probabilities are found in Table~\ref{tab:lucas_probs}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{|c|c|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{Conditional Probabilities}} \\
+        \hline
+        \textit{Conditional} & \textit{Probability} \\
+        \hline
+        P(Anxiety=T) & 0.64277 \\ \hline
+        P(Peer Pressure=T) & 0.32997 \\ \hline
+        P(Smoking=T $\vert$ Peer Pressure=F, Anxiety=F) & 0.43118 \\ \hline
+        P(Smoking=T $\vert$ Peer Pressure=T, Anxiety=F) & 0.74591 \\ \hline
+        P(Smoking=T $\vert$ Peer Pressure=F, Anxiety=T) & 0.8686 \\ \hline
+        P(Smoking=T $\vert$ Peer Pressure=T, Anxiety=T) & 0.91576 \\ \hline
+        P(Yellow Fingers=T $\vert$ Smoking=F) & 0.23119 \\ \hline
+        P(Yellow Fingers=T $\vert$ Smoking=T) & 0.95372 \\ \hline
+        P(Genetics=T) & 0.15953 \\ \hline
+        P(Lung cancer=T $\vert$ Genetics=F, Smoking=F) & 0.23146 \\ \hline
+        P(Lung cancer=T $\vert$ Genetics=T, Smoking=F) & 0.86996 \\ \hline
+        P(Lung cancer=T $\vert$ Genetics=F, Smoking=T) & 0.83934 \\ \hline
+        P(Lung cancer=T $\vert$ Genetics=T, Smoking=T) & 0.99351 \\ \hline
+        P(Attention Disorder=T $\vert$ Genetics=F) & 0.28956 \\ \hline
+        P(Attention Disorder=T $\vert$ Genetics=T) & 0.68706 \\ \hline
+        P(Born an Even Day=T) & 0.5 \\ \hline
+        P(Allergy=T) & 0.32841 \\ \hline
+        P(Coughing=T $\vert$ Allergy=F, Lung cancer=F) & 0.1347 \\ \hline
+        P(Coughing=T $\vert$ Allergy=T, Lung cancer=F) & 0.64592 \\ \hline
+        P(Coughing=T $\vert$ Allergy=F, Lung cancer=T) & 0.7664 \\ \hline
+        P(Coughing=T $\vert$ Allergy=T, Lung cancer=T) & 0.99947 \\ \hline
+        P(Fatigue=T $\vert$ Lung cancer=F, Coughing=F) & 0.35212 \\ \hline
+        P(Fatigue=T $\vert$ Lung cancer=T, Coughing=F) & 0.56514 \\ \hline
+        P(Fatigue=T $\vert$ Lung cancer=F, Coughing=T) & 0.80016 \\ \hline
+        P(Fatigue=T $\vert$ Lung cancer=T, Coughing=T) & 0.89589 \\ \hline
+        P(Car Accident=T $\vert$ Attention Disorder=F, Fatigue=F) & 0.2274 \\ \hline
+        P(Car Accident=T $\vert$ Attention Disorder=T, Fatigue=F) & 0.779 \\ \hline
+        P(Car Accident=T $\vert$ Attention Disorder=F, Fatigue=T) & 0.78861 \\ \hline
+        P(Car Accident=T $\vert$ Attention Disorder=T, Fatigue=T) & 0.97169 \\ \hline
+        \hline
+    \end{tabular}
+    \caption{Conditional probabilities for the nodes in the LUCAS0 dataset.}
+    \label{tab:lucas_probs}
+\end{table}
+
 \end{document}

### 2024-04-16 latex/draft.tex
Pushing latex changes.

@@
 \usetikzlibrary{shapes.geometric}
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
 \usepackage{hyperref}
@@
     P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{PP} = \text{T})), \\
-    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{A} = \text{T})), \\
+    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{A} = \text{T})), \numberthis \label{do_list}\\
     P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{AD} = \text{T})), \\
@@
 
-Plugging forward with \textbf{do} calculus, we can write $P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T}))$ as
-(using corollary~\ref{corollary:do})
+Plugging forward with \textbf{do} calculus, we can do one as an example. We write $P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T}))$
+(using corollary~\ref{corollary:do}) as
+
+\begin{equation}
+    P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T})) = 
+    \sum_{s\in\{T, F\}} \frac{P(\text{YF} = \text{T},\, \text{LC} = \text{T},\, \text{S} = s)}{P(\text{YF} = \text{T}\,|\, \text{S} = s)},
+    \label{yf1}
+\end{equation}
+
+where $\text{S}$ is smoking. Since the Lung Cancer node has Smoking and Genetics as parents, and Smoking
+has parent Anxiety and Peer Pressure (see Figure~\ref{fig:lucas_dag}), we can write Equation~\ref{yf1} to
+
+\begin{equation}
+    \sum_{S, G, A, PP}\frac{P(\text{YF} =\text{T}\;|\; \text{S}) P(\text{LC} 
+    = \text{T}\;|\; \text{S},\, \text{G}) P(\text{G}) P(\text{S}\;|\;\text{A},\, \text{PP}) P(\text{A})P(\text{PP})}
+    {P(\text{YF} = \text{T}\;|\; \text{S})}.
+    \label{yf2}
+\end{equation}
+
+where each $\text{S},\text{A}, \text{G}, \text{PP}$ is in terms of $\{T, F\}$.
+
+Since not every term is in terms of every variable, we can re-write the Equation~\ref{yf2} as
+
+\begin{align*}
+    &=\sum_{S} \frac{P(\text{YF} = T\,|\, S)}{P(\text{YF} = T\,|\, \text{S})}
+        \Biggl(\sum_{G} P(\text{LC} = \text{T}\,|\, \text{S}, \text{G})P(\text{G})\Biggr)
+            \Biggl(\sum_{A} P(\text{A}) 
+                \Bigl(\sum_{\text{PP}}P(\text{S}\,|\, \text{A}, \text{PP})P(\text{PP})\Bigr) 
+            \Biggr) \\
+    &=\sum_{S}\Big(\sum_{G} P(\text{LC} = \text{T}\,|\, \text{S}, \text{G})P(\text{G})\Big)
+        \Biggl(\sum_{A} P(\text{A}) 
+            \Bigl(\sum_{\text{PP}}P(\text{S}\,|\, \text{A}, \text{PP})P(\text{PP})\Bigr) 
+        \Biggr). \numberthis \label{yf3}
+\end{align*}
+
+Using the probabilities from Table~\ref{tab:lucas_probs} on Equation~\ref{yf3} we get
 
 \begin{equation*}
-    P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T})) = 
-    \sum_{s\in\{T, F\}} \frac{P(\text{YF} = \text{T},\, \text{LC} = \text{T},\, \text{S} = s)}{P(\text{YF} = \text{T}\,|\, \text{S} = s)},
+    P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T})) = {\color{red} \rule{2cm}{0.15mm} \text{ fix this!!}}
 \end{equation*}
 
-where $\text{S}$ is smoking. Using the property that we can split the numerator into probabilities only involving parents ({\color{red}rewrite})
+Performing similar calculations on the other probabilities in Equation~\ref{do_list}, we get
 
-\begin{equation*}
-    \sum_{s\in\{T, F\}}\frac{P(\text{YF} =\text{T}\;|\; \text{S} = s) P(\text{LC} 
-    = \text{T}\;|\; \text{S}=s,\, \text{G}) P(\text{S}=s\;|\;\text{A},\, \text{PP})}
-    {P(\text{YF} = \text{T}\;|\; \text{S} = s)}.
-\end{equation*}
+\begin{table}[h]
+    \centering
+    \begin{tabular}{|c|c|}
+        \hline
+        \textit{do Statements} & \textit{Probabilities} \\ \hline
+        $P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{YF} = \text{T})) &  \\ \hline
+        $P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{PP} = \text{T})) &  \\ \hline
+        $P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{A} = \text{T})) &   \\ \hline
+        $P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{AD} = \text{T})) &  \\ \hline
+        $P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{CA} = \text{T})) &  \\ \hline
+    \end{tabular}
+    \caption{The resulting probabilities of the \textbf{do} statements found in Equation~\ref{do_list}.
+    For a more detailed workthrough of these probabilities, see Appendix~\ref{do_work}.}
+    \label{tab:do_work_tab}
+\end{table}
 
-All of these fun conditionals lead us to
+% \begin{equation}
 
-\begin{equation*}
-    \sum_{s,a,g,p}\frac{P(\text{YF} =\text{T}\;|\; \text{S} = s) P(\text{LC} 
-    = \text{T}\;|\; \text{S}=s,\, \text{G}=g) P(\text{S}=s\;|\;\text{A}=a,\, \text{PP}=p)}
-    {P(\text{YF} = \text{T}\;|\; \text{S} = s)},
-\end{equation*}
-
-where each $s,a,g,p$ is summed over the set $\{T, F\}$.
+% \end{equation}
 
@@
 
+\section{Conditioning Outside Markov Blanket Results}
+\label{do_work}
+
+\subsection{$P$(\text{LC} = \text{T} \,$\vert$\, \textbf{do}(\text{PP} = \text{T}))}
+
+We have $P(\text{LC} = \text{T} \,\vert\, \textbf{do}(\text{PP} = \text{T}))$ yields
+
+\begin{align*}
+     &= P(\text{LC} = \text{T},\, \text{PP} = \text{T}) \\
+        &= P(\text{PP} = \text{T})\left(\sum_{s, g, an}P(\text{LC} = T \,|\, \text{S} = s,\, \text{G} = g)P(\text{G} = g)P(\text{An} = an)\right) \\
+        &= P(\text{PP} = \text{T})\Big(\sum_{an}P(\text{An} = an)\Big)\left(\sum_{s, g}P(\text{LC} = T \,|\, \text{S} = s,\, \text{G} = g)P(\text{G} = g)\right) \\
+        &= P(\text{PP} = \text{T})\left(\sum_{s, g}P(\text{LC} = T \,|\, \text{S} = s,\, \text{G} = g)P(\text{G} = g)\right) \\
+    &= \rule{2cm}{0.15mm}.
+\end{align*}
+
+\subsection{$P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{An} = \text{T}))}
+
+We have $P(\text{LC} = \text{T} \,\vert\, \textbf{do}(\text{An} = \text{T}))$ yields
+
+\begin{align*}
+     &= P(\text{LC} = \text{T},\, \text{An} = \text{T}) \\
+        &= P(\text{An} = \text{T})\left(\sum_{s, g, p}P(\text{LC} = T \,|\, \text{S} = s,\, \text{G} = g)P(\text{G} = g)P(\text{S}=s\,|\, \text{An} = T, \text{PP} = p)P(\text{PP} = p)\right) \\
+    &= \rule{2cm}{0.15mm}.
+\end{align*}
+
+\subsection{$P$(\text{LC} = \text{T} $\vert$ \textbf{do}(\text{AD} = \text{T}))}
+
+We have
+
+\begin{align*}
+    P(\text{LC} = \text{T} \,\vert\, \textbf{do}(\text{AD} = \text{T})) &= P(\text{LC} = \text{T},\, \text{AD} = \text{T},\, \text{G}) \\
+        &= \Biggl( \sum_{g} P(\text{AD} = \text{T}\,\vert\, \text{G} = g)\Bigl(\sum_{s}P(\text{LC} = T \,|\, \text{S} = s,\, \text{G} = g)\Bigr)\Biggr) \\
+    &= \rule{2cm}{0.15mm}.
+\end{align*}
+
 \end{document}

### 2024-04-16 latex/draft.tex
Small changes.

@@
 
-We have
+We have $P(\text{LC} = \text{T} \,\vert\, \textbf{do}(\text{AD} = \text{T}))$
 
 \begin{align*}
-    P(\text{LC} = \text{T} \,\vert\, \textbf{do}(\text{AD} = \text{T})) &= P(\text{LC} = \text{T},\, \text{AD} = \text{T},\, \text{G}) \\
-        &= \Biggl( \sum_{g} P(\text{AD} = \text{T}\,\vert\, \text{G} = g)\Bigl(\sum_{s}P(\text{LC} = T \,|\, \text{S} = s,\, \text{G} = g)\Bigr)\Biggr) \\
-    &= \rule{2cm}{0.15mm}.
+     &= \sum_{g} \frac{P(\text{LC} = \text{T},\, \text{AD} = \text{T},\, \text{G}=g)}{P(\text{AD} = T\;|\; G=g)} \\
+     &= \sum_{g, s, an,  p} \frac{P(\text{LC} = \text{T},\, \text{AD} = \text{T},\, \text{G}=g,\, \text{S} = s,\, \text{AN} = an,\, \text{PP} = p)}{P(\text{AD} = T\;|\; G=g)} \\
+    &= \sum_{g,s,an,p}\frac{P(\text{LC}=T\,|\,\text{S}=s, \text{G}=g)P(\text{G}=g)P(\text{S}=s\,|\,\text{AN}=an,\,\text{PP}=p)}{}.
 \end{align*}

### 2024-04-19 latex/draft.tex
intro and related work

@@
 \label{intro}
+Neural networks are at the center of computer science research and have revolutionized our ability to model complex relationships with data.
+However, as these networks become increasingly pervasive in decision-making processes, understanding the causality behind their predictions becomes paramount. 
+The distinction between correlation and causation is not merely philosophical; it's foundational for building robust, interpretable, and ethically sound AI systems.
+
+Causal inference, a field deeply rooted in statistics and philosophy, offers a framework to disentangle cause-and-effect relationships from observed data. 
+Traditionally, causal inference has been applied in fields such as epidemiology, economics, and social sciences. 
+However, its integration into the domain of neural networks presents a promising frontier, with implications ranging from improving model interpretability to enhancing the fairness and accountability of AI systems.
+
+In this paper, we explore the intersection of causal inference and neural networks, focusing on the application of Graph Neural Networks (GNNs) to causal modeling.
+We leverage the power of GNNs to learn representations of graph-structured data and perform a supervised classification task on the LUCAS0 dataset~\cite{lucas_dataset}.
+Our goal is to demonstrate the effectiveness of GNNs in capturing causal relationships within a graph and to provide insights into the interpretability and robustness of these models.
+
+\section{Related Work}
+\label{related}
+The integration of causal inference with neural networks has garnered significant interest in recent years.
+Several approaches have been proposed to combine the strengths of these two fields, with a particular focus on leveraging the expressive power of neural networks to model causal relationships.
+Koch et al. discusses ongoing work to extend causal inference to settings where confounding is non-linear, time-varying, or encoded in text, networks, and images \cite{koch2023deep}.
+Also, Yuan et al. compared the performance of CNN's on causal data to previous methods \cite{yuan2020causal}.
 
